### 2020-05-28 writeup/proof_examples.tex
B.1, B.2

Combining \eqref{hatw_add1} and \eqref{boundv-w}, we obtain that whp,
\be\label{hatv_add1} 
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2+\e} 
\ee
Inserting it into \eqref{eq_te_mtl_2task} and using a similar concentration result as in \eqref{boundv-w}, we get whp,
\begin{align}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\ 
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. With estimate \eqref{eq_isometric}, we see that 
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim \frac{p}{n_1+n_2}.$$
Hence we need to choose that $p\cdot d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+2\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the following choices of parameters: there exists a constant $c>0$ such that
\be\label{choiceofpara}
pd^2 \sim \sigma^2 \sim 1,\quad p^{-1+c}\kappa^2  \le d^2 \le p^c \kappa^2.
\ee
Under this choice, we can write 
\be\label{simpte1}
\begin{split}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  
\end{split}
\ee
whp for some constant $\e>0$.

With \eqref{simpte1} and Lemma \ref{prop_model_shift_tight}, we can prove Proposition \ref{prop_dist_transition}, which gives a transition threshold with respect to the ratio between the model bias and the noise level. With slight abuse of notations, we shall write $\hat a_i$, $\hat b_k$ and $\hat M$ as $a_i$, $b_k$ and $M$, respectively, throughout the rest of this section.

\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	In the setting of Proposition \ref{prop_dist_transition}, we have $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \id$. Then solving equations \eqref{eq_a2} and \eqref{eq_a3} with $\hat \lambda_i=1$, we get that
	\begin{align}
		 a_1 = \frac{\rho_1(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} , \label{simplesovlea12}\\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(\rho_1 + \rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(\rho_1 + \rho_2 - 1)}.\label{simplesovlea34}
	\end{align}
	Using Lemma \ref{lem_minv} and Lemma \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as 
$$\delta_{\vari}:=\sigma^2  \bigtr{(X_2^{\top}X_2)^{-1} }  - (1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } =\Delta_{\vari}\cdot (1+\OO(n^{-\e})) $$
with high probability, where 
	\begin{align}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{\rho_2 - 1} - \frac{1}{\rho_1 + \rho_2}\cdot\frac{1}{a_1+ a_2} } =\sigma^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)}.\label{Deltavar}
	\end{align}
	%where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next for the model shift bias
	$$\delta_\beta:=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right], $$
	we can get from Lemma \ref{prop_model_shift_tight} (or rather the proof of Lemma \ref{prop_model_shift_tight}) that
\be\label{Deltabeta} 
\al_-^2(\rho_1) - \oo(1)  \le \frac{\delta_\beta}{ \Delta_{\beta}} \le \al_+^2(\rho_1) +  \oo(1) , \ee
	where 
	$$\Delta_{\beta}:=pd^2 \cdot \frac{\rho_1^2}{(\rho_1+\rho_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2}= pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}.$$	
%\begin{align*}
%		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
%		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}d^2}
%	\end{align*}
%	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
%	\be\label{Deltabeta} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{p^{1/2+\e}d^2} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \ee	
%	Hence we obtain that
%	\begin{align*}
%		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
%		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
%	\end{align*}
With \eqref{Deltavar} and \eqref{Deltabeta}, we conclude that if
\be\label{upper101}pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \cdot \left( \al_+^2(\rho_1) +  \oo(1) \right) < \sigma^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)},\ee
	we have that $\delta_{\vari} > \delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$. We can simplify \eqref{upper101} to
	\[   d^2  < \frac{\sigma^2}{p}\frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1(\rho_1 + \rho_2) (\rho_2 - 1)} \cdot \left(\bigbrace{1 + \sqrt{\frac 1 {\rho_1}}}^{-4} - \oo(1)\right), \]
Plugging into $\rho_1>40$, we obtain the first statement of Proposition \ref{prop_dist_transition}.

@@

\begin{proof}[Proof of Proposition \ref{prop_var_transition}]
{\todo{rewrite the expressions for validation loss}} Under the setting of Proposition \ref{prop_var_transition}, we have
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \bignorm{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v} \beta_t)}^2 \nonumber \\
			&~ + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2)\hat v^2  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1}.
\end{align*}
Using concentration of random vector with i.i.d. Gaussian entries, Lemma \ref{largedeviation}, we obtain that
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \left[d^2 +\left( w-1\right)^2\kappa^2\right]\bigtr{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2} \cdot \left(1+\OO(p^{-1/2})\right)\nonumber \\
	& + (\sigma_1^2 -\sigma_2^2)\hat v^2  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }.
\end{align*}
In the setting of Proposition \ref{prop_var_transition}, we can also obtain that
$$|\hat w-1|=\OO(p^{-1}).$$
as in \eqref{hatw_add1}. We omit the details of the proof, since it is almost the same as the one in the proof of Lemma \ref{lem_hat_v}. Thus $\te(\hat{\beta}_t^{\MTL}) $ can be reduced to 
$$	\te(\hat{\beta}_t^{\MTL}) = \left[\Delta_\beta + \Delta_{diff}+  \sigma_2^2 \cdot \tr(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \right]\cdot \left(1+\OO(p^{-1/2})\right), $$
where $\Delta_\beta:= \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }$ has been estimated in \eqref{Deltabeta} and $\Delta_{diff}$ is
\begin{align*}	
	\Delta_{diff}:=(\sigma_1^2 -\sigma_2^2) \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{align*}
Using \eqref{eq_isometric} and Lemma \ref{lem_cov_derivative}, a similar bound holds for $\Delta_{diff}$ as in \eqref{Deltabeta}:
$$\bigbrace{1 - \sqrt{\frac{1}{c_1}}}^2\le \frac{\Delta_{diff}}{\sigma_1^2 -\sigma_2^2}\cdot \left( \frac{c_1 (c_1+c_2)}{(c_1+c_2-1)^3} + \OO(p^{-1/2+\e})\right)^{-1}\le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^2.$$

Next we compare $\Delta_{\var}$ with $\Delta_\beta + \Delta_{diff}$. Under the condition $d^2 < \frac {2\sigma^2} {3p} \frac{(c_1 + c_2 -1)^2}{c_1 (c_1 + c_2)(c_2 - 1)}$, if $\sigma_1^2\le \sigma_2^2$ then we always have that $\Delta_{\vari} > \Delta_{\beta}$ and hence $\te(\hat{\beta}_t^{\MTL})$ is smaller than $\te(\hat{\beta}_t^{\STL})$. It remains to consider the case $\sigma_1^2 \ge \sigma_2^2$. If 
\begin{align} 
&\sigma_2^2  \frac{c_1}{(c_2-1)(c_1 + c_2 -1)}\left(1 + \bigo{p^{-1/2 + \varepsilon}}\right) \label{cond sigma1}\\
&> \left[\frac{c_1^2 (c_1+c_2)}{(c_1+c_2-1)^3} + \bigo{p^{-1/2 + \varepsilon}}\right]\left[pd^2 \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 + \frac{\sigma_1^2-\sigma_2^2}{c_1} \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^2\right] ,\nonumber
\end{align}
then we have $\Delta_{\vari} > \Delta_{\beta}$ and we have $\te(\hat{\beta}_t^{\MTL})$ is smaller than $\te(\hat{\beta}_t^{\STL})$. For $c_1\ge 100$, we see that \eqref{cond sigma1} is implied by the following condition
$$\sigma_1^2 \le pd^2 \cdot c_1 + \left[ 1+ \frac23\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)} \right]\sigma_2^2$$
as long as $p$ is large enough. On the other hand, if 
\begin{align} 
&\sigma_2^2  \frac{c_1}{(c_2-1)(c_1 + c_2 -1)}\left(1 + \bigo{p^{-1/2 + \varepsilon}}\right) \label{cond sigma12}\\
&< \left[\frac{c_1^2 (c_1+c_2)}{(c_1+c_2-1)^3} + \bigo{p^{-1/2 + \varepsilon}}\right]\left[pd^2 \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 + \frac{\sigma_1^2-\sigma_2^2}{c_1} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^2\right] ,\nonumber
\end{align}
then we have $\Delta_{\vari} < \Delta_{\beta}$ and we have $\te(\hat{\beta}_t^{\MTL})$ is larger than $\te(\hat{\beta}_t^{\STL})$. For $c_1\ge 100$, we see that \eqref{cond sigma12} is implied by the following condition
$$\sigma_1^2 \ge  pd^2 \cdot c_1 + \left[ 1+ \frac32\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)} \right]\sigma_2^2.$$
\end{proof}

@@

@@

>>>>

Combining \eqref{hatw_add1} and \eqref{boundv-w}, we obtain that whp,
\be\label{hatv_add1} 
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2+\e} 
\ee
Inserting it into \eqref{eq_te_mtl_2task} and using a similar concentration result as in \eqref{boundv-w}, we get whp,
\begin{align}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\ 
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. With estimate \eqref{eq_isometric}, we see that 
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim \frac{p}{n_1+n_2}.$$
Hence we need to choose that $p\cdot d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+2\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the following choices of parameters: there exists a constant $c>0$ such that
\be\label{choiceofpara}
pd^2 \sim \sigma^2 \sim 1,\quad p^{-1+c}\kappa^2  \le d^2 \le p^c \kappa^2.
\ee
Under this choice, we can write 
\be\label{simple1}
\begin{split}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  
\end{split}
\ee
whp for some constant $\e>0$.

With \eqref{simple1} and Lemma \ref{prop_model_shift_tight}, we can prove Proposition \ref{prop_dist_transition}, which gives a transition threshold with respect to the ratio between the model bias and the noise level. With slight abuse of notations, we shall write $\hat a_i$, $\hat b_k$ and $\hat M$ as $a_i$, $b_k$ and $M$, respectively, throughout the rest of this section.

\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	In the setting of Proposition \ref{prop_dist_transition}, we have $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \id$. Then solving equations \eqref{eq_a2} and \eqref{eq_a3} with $\hat \lambda_i=1$, we get that
	\begin{align}
		 a_1 = \frac{\rho_1(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} , \label{simplesovlea12}\\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(\rho_1 + \rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(\rho_1 + \rho_2 - 1)}.\label{simplesovlea34}
	\end{align}
	Using Lemma \ref{lem_minv} and Lemma \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as 
\be\label{Deltavar}
\begin{split}
\delta_{\vari}&:=\sigma^2  \bigtr{(X_2^{\top}X_2)^{-1} }  - (1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \\
&=\Delta_{\vari}\cdot (1+\OO(n^{-\e})) 
\end{split}
\ee
with high probability, where 
	\begin{align*}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{\rho_2 - 1} - \frac{1}{\rho_1 + \rho_2}\cdot\frac{1}{a_1+ a_2} } =\sigma^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)}.
	\end{align*}
	%where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next for the model shift bias
	$$\delta_\beta:=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right], $$
	we can get from Lemma \ref{prop_model_shift_tight} (or rather the proof of Lemma \ref{prop_model_shift_tight}) that
\be\label{Deltabeta} 
\al_-^2(\rho_1) - \oo(1)  \le \frac{\delta_\beta}{ \Delta_{\beta}} \le \al_+^2(\rho_1) +  \oo(1) , \ee
	where 
	$$\Delta_{\beta}:=pd^2 \cdot \frac{\rho_1^2}{(\rho_1+\rho_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2}= pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}.$$	
%\begin{align*}
%		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
%		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}d^2}
%	\end{align*}
%	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
%	\be\label{Deltabeta} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{p^{1/2+\e}d^2} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \ee	
%	Hence we obtain that
%	\begin{align*}
%		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
%		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
%	\end{align*}
With \eqref{Deltavar} and \eqref{Deltabeta}, we conclude that if
\be\label{upper101}pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \cdot \left( \al_+^2(\rho_1) +  \oo(1) \right) < \sigma^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)},\ee
	we have that $\delta_{\vari} > \delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$. We can simplify \eqref{upper101} to
	\[   d^2  < \frac{\sigma^2}{p}\frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1(\rho_1 + \rho_2) (\rho_2 - 1)} \cdot \left(\bigbrace{1 + \sqrt{\frac 1 {\rho_1}}}^{-4} - \oo(1)\right), \]
Plugging into $\rho_1>40$, we obtain the first statement of Proposition \ref{prop_dist_transition}.

@@

\begin{proof}[Proof of Proposition \ref{prop_var_transition}]
In the setting of Proposition \ref{prop_var_transition}, the validation loss and the test error become
\begin{align*}
		\val(\hat{B}; w_1, w_2)
	&=  n_1 \cdot \bignorm{\Sigma_1^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_2^{\top}X_2\left (\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
		&+ n_1 \sigma^2 \cdot \frac{w_1^2}{w_2^2} \bigtr{\left(\frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)} \nonumber \\
		&+ n_2 \cdot \frac{w_1^2}{w_2^2}\bignorm{\Sigma_2^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_1^{\top}X_1\left(\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
		&+ n_2 \sigma^2 \cdot \bigtr{\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)}, 
\end{align*}
and 
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \bignorm{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v} \beta_t)}^2 \nonumber \\
			&~ + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2) \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} \hat v^2 X_1^\top X_1},
\end{align*}
where $\hat v=\hat{w_1}/\hat{w_2}$ is the global minimizer of $\val(\hat{B}; w_1, w_2)$. Again using concentration of random vector with i.i.d. entries, Lemma \ref{largedeviation}, we can rewrite $\te(\hat{\beta}_t^{\MTL})$ as
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \left[d^2 +\left( w-1\right)^2\kappa^2\right]\bigtr{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2} \cdot \left(1+\OO(p^{-1/2+\e})\right)\nonumber \\
	& + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2)  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} \hat v^2X_1^\top X_1}
\end{align*}
with high probability for any constant $\e>0$.

In the current setting, we can also show that  \eqref{hatv_add1}  holds for $\hat v$. 
Since the proof is almost the same as the one for Lemma \ref{lem_hat_v}, we omit the details. 
%$$|\hat w-1|=\OO(p^{-1}).$$
%as in \eqref{hatw_add1}. 
%We omit the details of the proof, since it is almost the same as the one in the proof of Lemma \ref{lem_hat_v}. 
Thus under the choice parameters in \eqref{choiceofpara}, $\te(\hat{\beta}_t^{\MTL}) $ can be simplified as in \eqref{simple1}: 
\be \label{simple2}
\begin{split}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \\
&+(1+\OO(n^{-\e}))\cdot (\sigma_1^2-\sigma_2^2)  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{split}
\ee
Then we write 
$$ \te(\hat{\beta}_t^{\STL})-\te(\hat{\beta}_t^{\MTL}) =\delta_{\vari} - \delta_\beta - \delta_{\vari}^{(2)},$$
where 
$$\delta_{\vari}:=\sigma_2^2  \bigtr{(X_2^{\top}X_2)^{-1} }  - (1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }$$
satisfies \eqref{Deltavar} but with $\sigma^2$ replaced with $\sigma_2^2$, 
$$\delta_\beta:=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]$$
satisfies \eqref{Deltabeta}, and 
\begin{align*}	
	\delta_{\vari}^{(2)}:=(1+\OO(n^{-\e}))\cdot (\sigma_1^2 -\sigma_2^2) \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{align*}
%$$	\te(\hat{\beta}_t^{\MTL}) = \left[\Delta_\beta + \Delta_{diff}+  \sigma_2^2 \cdot \tr(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \right]\cdot \left(1+\OO(p^{-1/2})\right), $$
%where $\Delta_\beta:= \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }$ has 
To estimate this new term, we use the same arguments as in the proof of Lemma \ref{prop_model_shift_tight}: we first replace $X_1^\top X_1$ with $n_1\id$ up to some error using \eqref{eq_isometric}, and then apply Lemma \ref{lem_cov_derivative} to calcualte $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2}}$. This process leads to the following estimates on $\delta_{\vari}^{(2)}$:
\be\label{Deltavar2} 
\al_-(\rho_1) - \oo(1)  \le \frac{\delta_{\vari}^{(2)}}{ \Delta_{\vari}^{(2)}} \le \al_+(\rho_1) +  \oo(1) , \ee
where 
$$ \Delta_{\vari}^{(2)}:=(\sigma_1^2 -\sigma_2^2) \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3}.$$
Next we compare $\delta_{\vari}$ with $\delta_\beta + \delta_{\vari}^{(2)}$. Our main goal is to see how the extra $\delta_{\vari}^{(2)}$ affects the information transfer in this case.

%Using \eqref{eq_isometric} and Lemma \ref{lem_cov_derivative}, a similar bound holds for $\Delta_{diff}$ as in \eqref{Deltabeta}:
%$$\bigbrace{1 - \sqrt{\frac{1}{c_1}}}^2\le \frac{\Delta_{diff}}{\sigma_1^2 -\sigma_2^2}\cdot \left( \frac{c_1 (c_1+c_2)}{(c_1+c_2-1)^3} + \OO(p^{-1/2+\e})\right)^{-1}\le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^2.$$

Note that the condition $d^2 < \frac {\sigma_2^2} {2p} \cdot \Phi(\rho_1, \rho_2)$ means the we have $\delta_{\vari}>\delta_{\beta} $ by Proposition \ref{prop_dist_transition}. Hence if $\sigma_1^2\le \sigma_2^2$, then $\delta_{\vari}^{(2)}<0$ and we always have $\delta_{\vari} > \delta_{\beta}+ \delta_{\vari}^{(2)}$, which gives $\te(\hat{\beta}_t^{\MTL})<\te(\hat{\beta}_t^{\STL})$.

It remains to consider the case $\sigma_1^2 \ge \sigma_2^2$.

\noindent{\bf Positive transfer.} By \eqref{Deltavar}, \eqref{Deltabeta} and \eqref{Deltavar2} above, if
\be\label{cond sigma1}
\begin{split} 
&\sigma_2^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)} \cdot (1-\oo(1)) \\
&>pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^4 + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^2 , 
\end{split}
\ee
then we have $\delta_{\vari} > \delta_{\beta} + \delta_{\vari}^{(2)}$ whp, which gives $\te(\hat{\beta}_t^{\MTL})<\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma1} to get
\begin{align*}
\sigma_1^2 < - pd^2 \cdot \rho_1\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^2 \cdot (1-\oo(1))+\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2)\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^{-2}\right]\cdot (1-\oo(1)).
\end{align*}
Plugging into $\rho_1>50$, we obtain the first claim of Proposition \ref{prop_var_transition} for positive transfer.

\noindent{\bf Negative transfer.} On the other hand, if 
\be\label{cond sigma12}
\begin{split} 
&\sigma_2^2 \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)}\cdot \left(1 + \oo(1)\right) \\
&< pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^4 + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^2 , 
\end{split}
\ee
then we have $\delta_{\vari} < \delta_{\beta} + \delta_{\vari}^{(2)}$ whp, which gives $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma12} to get
\begin{align*}
\sigma_1^2 > - pd^2 \cdot \rho_1\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^2 \cdot (1+\oo(1))+\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2)\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^{-2}\right]\cdot (1+\oo(1)).
\end{align*}
Plugging into $\rho_1>50$, we obtain the second claim of Proposition \ref{prop_var_transition} for negative transfer.
\end{proof}

@@

We first prove Proposition \ref{prop_data_size}, which describes the effect of source task data size on the information transfer.

@@

Then we prove Proposition \ref{prop_data_efficiency}, which gives precise upper and lower bounds on the data efficiency ratio.
